A Bias-Variance Decomposition of Information Transfer in Multi-Task Learning

### 2020-08-24 arxiv/setup.tex
sec 3 wip

>>>>

We make several standard assumptions on the covariance matrix $\Sigma_k$ and the random vector $z_i$, for $k = 1,\dots, t$ and $i = 1,\dots, n_k$.
First, for every entry of $z_i$, we assume that its fourth moment exists \todo{check}, that is, there exists a constant $C > 0$ such that
\begin{equation}\label{assmAhigh}
	\ex{z_{i,j}^4} \le C,\ 1\le j \le p.
\end{equation}
Second, we assume that the spectral norm of $\Sigma_k$ is bounded by a fixed constant, for every $k = 1,\dots, t$.
This is without loss of generality since we can always rescale $\Sigma_k$ so that its spectral norm satisfies the condition.

### 2020-08-24 arxiv/transfer.tex
sec 3 wip

As mentioned in Section \ref{sec_prelim}, the optimization objective \eqref{eq_mtl} is non-convex in general.
Therefore, we consider an arbitrary local minimum $B, W_1, W_2$, in particular when $B = \hat{B}(W_1, W_2)$ satisfies the local optimality condition.
We derive deterministic conditions to compare the prediction loss of the local minimum to single-task learning.
Our key insight is a bias-variance decomposition of the expected prediction loss of $\hat{\beta}_2^{\MTL} = \hat{B} W_2$ as follows
\begin{align}
	\exarg{\epsilon_1, \epsilon_2}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2}
	=&~ \frac{W_1^2}{W_2^2} \bignorm{\Sigma_2^{1/2}(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_1 - \frac{W_1}{W_2} \beta_2)}^2 \label{eq_bias_2task} \\
			&+~  \sigma^2\cdot \bigtr{\Sigma_2(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} }. \label{eq_var_2task}
\end{align}
See Section \ref{sec_proof_general} for its derivation.
Equation \eqref{eq_bias_2task} is the bias of $\hat{\beta}_t^{\MTL}$ and
equation \eqref{eq_var_2task} is the variance of $\hat{\beta}_t^{\MTL}$.
%minus the variance of $\hat{\beta}_t^{\STL}$, which is always negative.
Comparing the above to single-task learning, that is,
\begin{align}
	\exarg{\epsilon_2}{L(\hat{\beta}_2^{\STL}) \mid X_2} = \sigma^2 \cdot \bigtr{\Sigma_2 (X_2^{\top} X_2)^{-1}}, \label{eq_var_stl}
\end{align}
we observe that while the bias of $\hat{\beta}_2^{\MTL}$ is always larger than that of $\hat{\beta}_2^{\STL}$, which is zero, the variance of $\hat{\beta}_2^{\MTL}$ is always lower than that of $\hat{\beta}_2^{\STL}$.\footnote{To see why this is true, we apply the Woodbury matrix identity over equation \eqref{eq_var_2task} and use the fact that for the product of two PSD matrices, its trace is always nonnegative.}
%As an example, for the setting of two tasks, we can decompose $L(\hat{\beta}_t^{\MTL}) - L(\hat{\beta}_t^{\STL})$ into a bias term and a variance term as follows .
%Recall that $\hat{\beta}_t^{\MTL}$ is defined as $BW_t$ after solving equation \eqref{eq_mtl}.
%We can  the test error of $\hat{\beta}_{t}^{\MTL}$ on the target task into two parts as follows.
%{\small\begin{align}
%	\te(\hat{\beta}_t^{\MTL}) - L(\hat{\beta}_t^{\STL}) =& ~ \hat{v}^2 \bignorm{\Sigma_2^{1/2} (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1 (\beta_1 - \hat{v}\beta_2)}^2 \label{eq_te_model_shift} \\
%	+&~ \sigma^2 \bigbrace{\bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} - \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2}}. \label{eq_te_var}
%\end{align}}%
%In the above, $\hat{v} = W_1 / W_2$ where $W_1, W_2$ are obtained from solving equation \eqref{eq_mtl_eval} (recalling that $W_1, W_2$ are scalars for two tasks).
%The role of $\hat{v}$ is to scale the shared subspace $B$ to fit each task.
Hence, the bias term introduces a negative effect that depends on the \textit{similarity} between $\beta_1$ and $\beta_2$.
Intuitively, the more \textit{samples} we have, the smaller the variance is.
Meanwhile, \textit{covariate shift} also affects how small the variance can be.

@@

We now state our main result for two tasks with both covariate and model shift. It shows that the information transfer is determined by two deterministic quantities $\Delta_{\bias}$ and $\Delta_{\vari}$, which give the change of model shift bias and the change of variance, respectively. The exact forms of $\Delta_{\bias}$ and $\Delta_{\vari}$ will be given in Lemma \ref{thm_model_shift}.

\begin{theorem}[Two tasks]\label{thm_main_informal}
	For the setting of two tasks, let $\delta > 0$ be a fixed error margin, $\rho_2 > 1$ and $\rho_1 \gtrsim \delta^{-2}\cdot \lambda_{\min}(\Sigma_1^{1/2}\Sigma_2^{-1/2})^{-4} \norm{\Sigma_1} \max(\norm{\beta_1}^2, \norm{\beta_2}^2)$.
 	There exist two deterministic functions $\Delta_{\bias}$ and $\Delta_{\vari}$ that only depend on $\set{\hat{v}, \Sigma_1, \Sigma_2, \rho_1, \rho_2, \beta_1, \beta_2}$ such that
	\squishlist
		\item If $\Delta_{\bias} - \Delta_{\vari} < -\delta$, then w.h.p. over the randomness of $X_1, X_2$, we have $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\Delta_{\bias} - \Delta_{\vari} > \delta$, then w.h.p. over the randomness of $X_1, X_2$, we have $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{theorem}

@@

%\textbf{Proof overview.}\todo{}
%Theorem \ref{lem_cov_shift_informal} extends a well-known result for the single-task setting when $X_1, \rho_1, a_1$ are all equal to zero \cite{S07}.
%Applying Theorem \ref{lem_cov_shift_informal} to \eqref{eq_te_var}, we can calculate the amount of reduced variance compared to STL, which is given asymptotically by $\Delta_{\vari}$.
%For the bias term in equation \eqref{eq_te_model_shift}, we apply the approximate isometry property to $X_1^{\top}X_1$, which is close to $n_1^2\Sigma_1$. This results in the error term $\delta$, which scales as $(1 + 1/\sqrt{\rho_1})^4-1$.
%Then, we apply a similar identity to Theorem \ref{lem_cov_shift_informal} for bounding the bias term, noting that the derivative of $R(z)$ with respect to $z$ can be approximated by $R_\infty'(z)$.
%This estimates the negative effect given by $\Delta_{\bias}$. %, which will be used to estimate the first term on the right hand side of \eqref{eq_te_model_shift}.
%During this process, we will get the $\Delta_{\bias}$ term up to an error $\delta$ depending on $\rho_1$.
The proof of Theorem \ref{thm_main_informal} is presented in Appendix \ref{app_proof_main_thm} and the proof of Lemma \ref{lem_cov_shift_informal} is in Appendix \ref{sec_maintools}.

%The formal statement is stated in Theorem \ref{thm_many_tasks} and its proof can be found in Appendix \ref{app_proof_many_tasks}.
%The technical crux of our approach is to derive the asymptotic limit of $\te(\hat{\beta}_t^{\MTL})$ in the high-dimensional setting, when $p$ approaches infinity.
%We derive a precise limit of $\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}$, which is a deterministic function that only depends on $\Sigma_1, \Sigma_2$ and $n_1/p, n_2/p$ (see Lemma \ref{lem_cov_shift} in Appendix \ref{app_proof_main} for the result).
%Based on the result, we show how to determine positive versus negative transfer as follows.
%, where $\lambda_{\min}(M)$ is the smallest singular value of $M_1$

Next, we derive a closed-form solution of the multi-task learning estimator for the case of two tasks.
From \cite{WZR20}, we know that we need to explicitly restrict the output dimension $r$ of $B$ so that there is transfer between the two tasks.
Hence for the case of two tasks, we consider the setting where $r=1$.
For simplicity of notations, we shall denote $(X_i^{tr},Y_i^{tr})$ and $(X_i^{val},Y_i^{val})$ as $(X_i,Y_i)$ and  $(\wt X_i,\wt Y_i)$, respectively. Then equation \eqref{eq_mtl} simplifies to
\begin{align}\label{eq_mtl_2task}
	f(B; w_1, w_2) = \bignorm{X_1 B w_1 - Y_1}^2 + \bignorm{X_2 B w_2 - Y_2}^2,
\end{align}
where $B\in\real^p$ and $w_1, w_2$ are both real numbers. To solve the above problem, suppose that $w_1, w_2$ are fixed, by local optimality, we find the optimal $B$ as
\begin{align}
	& \hat{B}(w_1, w_2) = (w_1^2 X_1^{\top}X_1 + w_2^2 X_2^{\top}X_2)^{-1} (w_1 X_1^{\top}Y_1 + w_2 X_2^{\top}Y_2) \label{hatB}\\
	&= \frac{1}{w_2} \left( \frac{w_1^2}{w_2^2}  X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} \left(\frac{w_1}{w_2} X_1^{\top}Y_1 + X_2^{\top}Y_2\right) \nonumber\\
	&= \frac{1}{w_2}\left[\beta_2 + \left(\frac{w_1^2}{w_2^2} X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}\bigbrace{X_1^{\top}X_1\left(\frac{w_1}{w_2}\beta_1 - \frac{w_1^2}{w_2^2} \beta_2\right) + \left(\frac{w_1}{w_2} X_1^{\top}\varepsilon_1 + X_2^{\top}\varepsilon_2\right)}\right]. \nonumber
\end{align}
As a remark, when $w_1 = w_2 = 1$, we obtain linear regression.
If $\beta_1$ is a scaling of $\beta_2$, then  $w_1, w_2$ can be scaled accordingly to fix both tasks more accurately than linear regression.

@@

\iffalse
{\color{red}\begin{align*} 
			& f(W_1, W_2) = \bignorm{X_1 \hat B w_1 - Y_1}^2 + \bignorm{X_2 \hat B w_2 - Y_2}^2\\
			& =\bignorm{X_1\left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} \left(v^2 X_1^{\top}Y_1 + vX_2^{\top}Y_2\right) - Y_1}^2 \\
			&+ \bignorm{X_2 \left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} \left(vX_1^{\top}Y_1 + X_2^{\top}Y_2\right) - Y_2}^2 \\
			& =\bignorm{X_1\left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} \left(v^2 X_1^{\top}\e_1 + vX_2^{\top}\e_2\right) - \e_1 + X_1\left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}  X_2^{\top}X_2(v\beta_2-\beta_1) }^2 \\
			&+ \bignorm{X_2 \left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} \left(vX_1^{\top}\e_1 + X_2^{\top}\e_2\right) - \e_2 + vX_2 \left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}  X_1^{\top}X_1(\beta_1-v\beta_2) }^2 \\
			&=\val(v)\cdot \left( 1+\OO(p^{-1/2\e})\right) \quad \text{w.h.p.},
		\end{align*}
		where
		\begin{align*}
		\val(v)&=\bignorm{X_1\left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}  X_2^{\top}X_2(v\beta_2-\beta_1) }^2 \\
			&+ v^2\bignorm{X_2 \left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}  X_1^{\top}X_1(\beta_1-v\beta_2) }^2 \\
			&+\sigma^2 \tr\left(v^2 X_1\left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}   X_1^{\top} -\id\right)^2 \\
			&+\sigma^2 \tr\left(v^2 X_1\left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}X_2^{\top}X_2 \left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} X_1^\top\right)\\
			&+\sigma^2 \tr\left(X_2 \left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} X_2^{\top} -\id\right)^2 \\
			&+\sigma^2 \tr\left(v^2 X_2\left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}X_1^{\top}X_1 \left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} X_2^\top\right)\\
			&=\bignorm{X_1\left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}  X_2^{\top}X_2(v\beta_2-\beta_1) }^2 \\
			&+ v^2\bignorm{X_2 \left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}  X_1^{\top}X_1(\beta_1-v\beta_2) }^2 +(n_1+n_2-p)\sigma^2.
		\end{align*}
	}
\fi

@@

Hence to minimize $g(v)$, it suffices to minimize $\val(v)$ over $v$.
Let $\hat v=\hat{w_1}/\hat{w_2}$ be the global minimizer of $g(v)$.
Now we can define the multi-task learning estimator for the target task as
	\[ \hat{\beta}_2^{\MTL} = \hat{w}_{2}\hat{B}(\hat{w}_1, \hat{w}_2) .\]
%	where $t=2$ since we are considering the two task case, and it also stands for the ``target task".
%The intuition for deriving $\hat{\beta}_2^{\MTL}$ is akin to performing multi-task training in practice.
%Let $\hat{v} = \hat{w_1} / \hat{w_2}$ for the simplicity of notation.
The prediction loss of using $\hat{\beta}_2^{\MTL}$ for the target task is
\begin{align}
	\te(\hat{\beta}_2^{\MTL}) =&~ \hat{v}^2 \bignorm{\Sigma_2^{1/2}(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_1 - \hat{v} \beta_2)}^2 \nonumber \\
			&+~  \bigtr{\Sigma_2(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-2}\left(\sigma_1^2 \cdot \hat v^2  X_1^{\top}X_1 + \sigma_2^2  \cdot X_2^{\top}X_2\right) }, \label{eq_te_mtl_2task}
\end{align}
which only depends on $\hat v$, the sample covariance matrices, and $\beta_1,\beta_2$.

@@

We shall refer to random matrices of the form $X^\top X$ as sample covariance matrices following the standard notations in high-dimensional statistics. The following two lemmas  are the main random matrix theoretical results of this paper, which will be used to estimate the two terms on the right-hand side of \eqref{eq_te_mtl_2task}.
%which deals with the inverse of the sum of two random matrices, which
%any is can be viewed as a special case of Theorem \ref{thm_model_shift}.

>>>>

As mentioned in Section \ref{sec_prelim}, the optimization objective \eqref{eq_mtl} is non-convex in general.
Therefore, we consider an arbitrary local minimum $B, W_1, W_2$, in particular when $B = \hat{B}(W_1, W_2)$ satisfies the local optimality condition.
We derive deterministic conditions to compare the prediction loss of the local minimum to single-task learning.
Our key insight is a bias-variance decomposition of the expected prediction loss of $\hat{\beta}_2^{\MTL} = \hat{B} W_2$ as follows
\begin{align}
	\exarg{\epsilon_1, \epsilon_2}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2}
	=&~ \frac{W_1^2}{W_2^2} \bignorm{\Sigma_2^{1/2}(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_1 - \frac{W_1}{W_2} \beta_2)}^2 \label{eq_bias_2task} \\
			&+~  \sigma^2\cdot \bigtr{\Sigma_2(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} }. \label{eq_var_2task}
\end{align}
See Section \ref{sec_proof_general} for its derivation.
Equation \eqref{eq_bias_2task} is the bias of $\hat{\beta}_t^{\MTL}$ and
equation \eqref{eq_var_2task} is the variance of $\hat{\beta}_t^{\MTL}$.
%minus the variance of $\hat{\beta}_t^{\STL}$, which is always negative.
Comparing the above to single-task learning, that is,
\begin{align}
	\exarg{\epsilon_2}{L(\hat{\beta}_2^{\STL}) \mid X_2} = \sigma^2 \cdot \bigtr{\Sigma_2 (X_2^{\top} X_2)^{-1}}, \label{eq_var_stl}
\end{align}
we observe that while the bias of $\hat{\beta}_2^{\MTL}$ is always larger than that of $\hat{\beta}_2^{\STL}$, which is zero, the variance of $\hat{\beta}_2^{\MTL}$ is always lower than that of $\hat{\beta}_2^{\STL}$.\footnote{To see why this is true, we apply the Woodbury matrix identity over equation \eqref{eq_var_2task} and use the fact that for the product of two PSD matrices, its trace is always nonnegative.}
%As an example, for the setting of two tasks, we can decompose $L(\hat{\beta}_t^{\MTL}) - L(\hat{\beta}_t^{\STL})$ into a bias term and a variance term as follows .
%Recall that $\hat{\beta}_t^{\MTL}$ is defined as $BW_t$ after solving equation \eqref{eq_mtl}.
%We can  the test error of $\hat{\beta}_{t}^{\MTL}$ on the target task into two parts as follows.
%{\small\begin{align}
%	\te(\hat{\beta}_t^{\MTL}) - L(\hat{\beta}_t^{\STL}) =& ~ \hat{v}^2 \bignorm{\Sigma_2^{1/2} (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1 (\beta_1 - \hat{v}\beta_2)}^2 \label{eq_te_model_shift} \\
%	+&~ \sigma^2 \bigbrace{\bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} - \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2}}. \label{eq_te_var}
%\end{align}}%
%In the above, $\hat{v} = W_1 / W_2$ where $W_1, W_2$ are obtained from solving equation \eqref{eq_mtl_eval} (recalling that $W_1, W_2$ are scalars for two tasks).
%The role of $\hat{v}$ is to scale the shared subspace $B$ to fit each task.
In other words, training both tasks together helps predict the target task by reducing variance while incurring a bias.
Therefore, whether multi-task learning outperforms single-task learning is determined by the bias-variance decomposition!
%Hence, the bias term introduces a negative effect that depends on the \textit{similarity} between $\beta_1$ and $\beta_2$.
%Intuitively, the more \textit{samples} we have, the smaller the variance is.
%Meanwhile, \textit{covariate shift} also affects how small the variance can be.

@@

Based on the above intuition, in the following we develop generalization bounds that relate the bias and variance of multi-task learning to the sample sizes and the covariance matrices of both tasks.

%We now state our main result for two tasks with both covariate and model shift. It shows that the information transfer is determined by two deterministic quantities $\Delta_{\bias}$ and $\Delta_{\vari}$, which give the change of model shift bias and the change of variance, respectively. The exact forms of $\Delta_{\bias}$ and $\Delta_{\vari}$ will be given in Lemma \ref{thm_model_shift}.

\begin{theorem}[Two-task case]\label{thm_main_informal}
	For the setting of two tasks, let $\delta > 0$ be a fixed error margin, $\rho_2 > 1$ and $\rho_1 \gtrsim \delta^{-2}\cdot \lambda_{\min}(\Sigma_1^{1/2}\Sigma_2^{-1/2})^{-4} \norm{\Sigma_1} \max(\norm{\beta_1}^2, \norm{\beta_2}^2)$.
	Let $B, W_1, W_2$ be any local minimum of equation \eqref{eq_mtl}.
 	There exist two deterministic functions $\Delta_{\bias}$ and $\Delta_{\vari}$ that only depend on $\set{W_1/W_2, \Sigma_1, \Sigma_2, \rho_1, \rho_2, \beta_1, \beta_2}$ such that
	\squishlist
		\item If $\Delta_{\bias} - \Delta_{\vari} < -\delta$, then w.h.p. over the randomness of $X_1, X_2$, we have $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\Delta_{\bias} - \Delta_{\vari} > \delta$, then w.h.p. over the randomness of $X_1, X_2$, we have $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{theorem}

@@

%\textbf{Proof overview.}\todo{}
%Theorem \ref{lem_cov_shift_informal} extends a well-known result for the single-task setting when $X_1, \rho_1, a_1$ are all equal to zero \cite{S07}.
%Applying Theorem \ref{lem_cov_shift_informal} to \eqref{eq_te_var}, we can calculate the amount of reduced variance compared to STL, which is given asymptotically by $\Delta_{\vari}$.
%For the bias term in equation \eqref{eq_te_model_shift}, we apply the approximate isometry property to $X_1^{\top}X_1$, which is close to $n_1^2\Sigma_1$. This results in the error term $\delta$, which scales as $(1 + 1/\sqrt{\rho_1})^4-1$.
%Then, we apply a similar identity to Theorem \ref{lem_cov_shift_informal} for bounding the bias term, noting that the derivative of $R(z)$ with respect to $z$ can be approximated by $R_\infty'(z)$.
%This estimates the negative effect given by $\Delta_{\bias}$. %, which will be used to estimate the first term on the right hand side of \eqref{eq_te_model_shift}.
%During this process, we will get the $\Delta_{\bias}$ term up to an error $\delta$ depending on $\rho_1$.
%The proof of Theorem \ref{thm_main_informal} is presented in Appendix \ref{app_proof_main_thm} and the proof of Lemma \ref{lem_cov_shift_informal} is in Appendix \ref{sec_maintools}.
%
%
%
%The formal statement is stated in Theorem \ref{thm_many_tasks} and its proof can be found in Appendix \ref{app_proof_many_tasks}.
%The technical crux of our approach is to derive the asymptotic limit of $\te(\hat{\beta}_t^{\MTL})$ in the high-dimensional setting, when $p$ approaches infinity.
%We derive a precise limit of $\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}$, which is a deterministic function that only depends on $\Sigma_1, \Sigma_2$ and $n_1/p, n_2/p$ (see Lemma \ref{lem_cov_shift} in Appendix \ref{app_proof_main} for the result).
%Based on the result, we show how to determine positive versus negative transfer as follows.
%, where $\lambda_{\min}(M)$ is the smallest singular value of $M_1$

\iffalse
Next, we derive a closed-form solution of the multi-task learning estimator for the case of two tasks.
From \cite{WZR20}, we know that we need to explicitly restrict the output dimension $r$ of $B$ so that there is transfer between the two tasks.
Hence for the case of two tasks, we consider the setting where $r=1$.
For simplicity of notations, we shall denote $(X_i^{tr},Y_i^{tr})$ and $(X_i^{val},Y_i^{val})$ as $(X_i,Y_i)$ and  $(\wt X_i,\wt Y_i)$, respectively. Then equation \eqref{eq_mtl} simplifies to
\begin{align}\label{eq_mtl_2task}
	f(B; w_1, w_2) = \bignorm{X_1 B w_1 - Y_1}^2 + \bignorm{X_2 B w_2 - Y_2}^2,
\end{align}
where $B\in\real^p$ and $w_1, w_2$ are both real numbers. To solve the above problem, suppose that $w_1, w_2$ are fixed, by local optimality, we find the optimal $B$ as
\begin{align}
	& \hat{B}(w_1, w_2) = (w_1^2 X_1^{\top}X_1 + w_2^2 X_2^{\top}X_2)^{-1} (w_1 X_1^{\top}Y_1 + w_2 X_2^{\top}Y_2) \label{hatB}\\
	&= \frac{1}{w_2} \left( \frac{w_1^2}{w_2^2}  X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} \left(\frac{w_1}{w_2} X_1^{\top}Y_1 + X_2^{\top}Y_2\right) \nonumber\\
	&= \frac{1}{w_2}\left[\beta_2 + \left(\frac{w_1^2}{w_2^2} X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}\bigbrace{X_1^{\top}X_1\left(\frac{w_1}{w_2}\beta_1 - \frac{w_1^2}{w_2^2} \beta_2\right) + \left(\frac{w_1}{w_2} X_1^{\top}\varepsilon_1 + X_2^{\top}\varepsilon_2\right)}\right]. \nonumber
\end{align}
As a remark, when $w_1 = w_2 = 1$, we obtain linear regression.
If $\beta_1$ is a scaling of $\beta_2$, then  $w_1, w_2$ can be scaled accordingly to fix both tasks more accurately than linear regression.

@@

{\color{red}\begin{align*}
			& f(W_1, W_2) = \bignorm{X_1 \hat B w_1 - Y_1}^2 + \bignorm{X_2 \hat B w_2 - Y_2}^2\\
			& =\bignorm{X_1\left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} \left(v^2 X_1^{\top}Y_1 + vX_2^{\top}Y_2\right) - Y_1}^2 \\
			&+ \bignorm{X_2 \left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} \left(vX_1^{\top}Y_1 + X_2^{\top}Y_2\right) - Y_2}^2 \\
			& =\bignorm{X_1\left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} \left(v^2 X_1^{\top}\e_1 + vX_2^{\top}\e_2\right) - \e_1 + X_1\left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}  X_2^{\top}X_2(v\beta_2-\beta_1) }^2 \\
			&+ \bignorm{X_2 \left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} \left(vX_1^{\top}\e_1 + X_2^{\top}\e_2\right) - \e_2 + vX_2 \left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}  X_1^{\top}X_1(\beta_1-v\beta_2) }^2 \\
			&=\val(v)\cdot \left( 1+\OO(p^{-1/2\e})\right) \quad \text{w.h.p.},
		\end{align*}
		where
		\begin{align*}
		\val(v)&=\bignorm{X_1\left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}  X_2^{\top}X_2(v\beta_2-\beta_1) }^2 \\
			&+ v^2\bignorm{X_2 \left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}  X_1^{\top}X_1(\beta_1-v\beta_2) }^2 \\
			&+\sigma^2 \tr\left(v^2 X_1\left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}   X_1^{\top} -\id\right)^2 \\
			&+\sigma^2 \tr\left(v^2 X_1\left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}X_2^{\top}X_2 \left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} X_1^\top\right)\\
			&+\sigma^2 \tr\left(X_2 \left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} X_2^{\top} -\id\right)^2 \\
			&+\sigma^2 \tr\left(v^2 X_2\left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}X_1^{\top}X_1 \left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} X_2^\top\right)\\
			&=\bignorm{X_1\left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}  X_2^{\top}X_2(v\beta_2-\beta_1) }^2 \\
			&+ v^2\bignorm{X_2 \left( v^2 X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}  X_1^{\top}X_1(\beta_1-v\beta_2) }^2 +(n_1+n_2-p)\sigma^2.
		\end{align*}
	}

@@

Hence to minimize $g(v)$, it suffices to minimize $\val(v)$ over $v$.
Let $\hat v=\hat{w_1}/\hat{w_2}$ be the global minimizer of $g(v)$.
Now we can define the multi-task learning estimator for the target task as
	\[ \hat{\beta}_2^{\MTL} = \hat{w}_{2}\hat{B}(\hat{w}_1, \hat{w}_2) .\]
%	where $t=2$ since we are considering the two task case, and it also stands for the ``target task".
%The intuition for deriving $\hat{\beta}_2^{\MTL}$ is akin to performing multi-task training in practice.
%Let $\hat{v} = \hat{w_1} / \hat{w_2}$ for the simplicity of notation.
The prediction loss of using $\hat{\beta}_2^{\MTL}$ for the target task is
\begin{align}
	\te(\hat{\beta}_2^{\MTL}) =&~ \hat{v}^2 \bignorm{\Sigma_2^{1/2}(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_1 - \hat{v} \beta_2)}^2 \nonumber \\
			&+~  \bigtr{\Sigma_2(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-2}\left(\sigma_1^2 \cdot \hat v^2  X_1^{\top}X_1 + \sigma_2^2  \cdot X_2^{\top}X_2\right) }, \label{eq_te_mtl_2task}
\end{align}
which only depends on $\hat v$, the sample covariance matrices, and $\beta_1,\beta_2$.
\fi

@@

\paragraph{Proof overview.}
We shall refer to random matrices of the form $X^\top X$ as sample covariance matrices following the standard notations in high-dimensional statistics. The following two lemmas  are the main random matrix theoretical results of this paper, which will be used to estimate the two terms on the right-hand side of \eqref{eq_te_mtl_2task}.
%which deals with the inverse of the sum of two random matrices, which
%any is can be viewed as a special case of Theorem \ref{thm_model_shift}.
